---
title: "Sales Data Explore"
author: "Charles Frenzel"
date: "Tuesday, January 26, 2015"
output:
  html_document:
    theme: readable
---

```{r    echo=FALSE, warning=FALSE, message=FALSE}
rm(list = ls(all = TRUE))

library(RODBC)
library(plyr)
library(dplyr)
library(ggvis)
library(ggplot2)
library(reshape2)
```


```{r, echo=T, message=FALSE, eval=FALSE , warning=FALSE}
load(mth, file="monthly_sales.RData")
load(brnd, file="brands.RData")
```


Merge back the data and look at some brief summaries

```{r, echo=FALSE, warning=FALSE, message=FALSE}
df <- left_join(mth, brnd, by = c("BRNDID"))
df <- tbl_df(df)

df$BRND_NM <- toupper(df$BRND_NM)
df$Date <- as.Date(paste(df$Year,df$Month,"1",sep="-"))
df <- df %>%  arrange(Date)
names(df) <- tolower(names(df))


p_tbl <- df %>% 
     filter(year == 2014, month != 11) %>%
     select(amt, cnt, id, month, brnd_nm) %>%
     mutate(Period =  ifelse(month < 6 , "First_5_Months", "Last_5_Months")) %>%
    group_by(Period, brnd_nm) %>%
     summarize(Total_Amounts = round(sum(amt, na.rm=T),0)
                ,Total_Counts = round(sum(cnt, na.rm=T),0)
                ,Unique_Custs = n_distinct(id)
                ,Count_of_Custs = n() ) %>%
  select(Total_Amounts, Total_Counts, Unique_Custs, Count_of_Custs, Brand_Name = brnd_nm, Period) %>%
  arrange(desc(Unique_Custs))
```


####Change in Amounts Frist 5 Months vs Last 5 Months

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9}
amt_tbl = p_tbl  %>%
        select(Period, Brand_Name, Total_Amounts) %>%
         dcast(., Brand_Name ~ Period) %>%
          mutate( Change = paste(round( ((Last_5_Months - First_5_Months) / First_5_Months), 2)*100,"%",sep="")
                  ,Change_num = round( ((Last_5_Months - First_5_Months) / First_5_Months), 2) ) %>%
            arrange(desc(Change_num))

ggplot(data=amt_tbl, aes(x=Brand_Name, y=Change_num), fill = "orange") + 
    geom_bar(colour="black", stat="identity", position=position_dodge(),  size=.3, fill ="blue") + 
    geom_text(aes(label = Change, y=Change_num + 0.05), size = 6) +
    xlab("Brand Name") + ylab("Percent Change") + # Set axis labels
    ggtitle("Percent Change First 5 Months vs Last 5 Months") +  # Set title
    theme_bw() + theme(axis.ticks = element_blank(), axis.text.y = element_blank())
rm(am_tbl)
```


#### Change in Counts Frist 5 Months vs Last 5 Months

```{r, echo=FALSE, warning=FALSE, message=FALSE,  fig.width=9}
cnt_tbl = p_tbl  %>%
        select(Period, Brand_Name, Total_Counts) %>%
         dcast(., Brand_Name ~ Period) %>%
          mutate( Change = paste(round( ((Last_5_Months - First_5_Months) / First_5_Months), 2)*100,"%",sep="")
                  ,Change_num = round( ((Last_5_Months - First_5_Months) / First_5_Months), 2) ) %>%
            arrange(desc(Change_num))

ggplot(data=cnt_tbl, aes(x=Brand_Name, y=Change_num)) + 
    geom_bar(colour="black", stat="identity", position=position_dodge(),  size=.3, fill ="orange") + 
    geom_text(aes(label = Change, y=Change_num + 0.05), size = 6) +
    xlab("Brand Name") + ylab("Percent Change") + # Set axis labels
    ggtitle("Percent Change First 5 Months vs Last 5 Months") +  # Set title
    theme_bw() + theme(axis.ticks = element_blank(), axis.text.y = element_blank())
rm(cnt_tbl)
```


#### Trend Amounts over Time

``` {r , echo=FALSE  , warning=FALSE,  results = "asis", fig.align='center'}
  df %>% 
  filter( brnd_nm == "T") %>%
    select(Brand_Name = brnd_nm, date,  amt) %>% 
     arrange(date) %>%
      group_by (date) %>%
         summarize(Amounts = sum(amt, na.rm=T)) %>%
  ggvis( ~as.factor(date) , ~Amounts ) %>% layer_lines( stroke :="#FF9900" , strokeWidth := 4 ) %>% 
  add_axis("x", title="Period", title_offset = 50 ,  properties = axis_props(labels = list( fontSize = 8))) %>% 
  add_axis("y", title="") %>%
  scale_numeric("y", domain = c(0,  120050000)) %>%
   add_axis("x", orient = "top", ticks = 0, title = "T Amounts By Month",
           properties = axis_props(
             axis = list(stroke = "white"),
             labels = list(fontSize = 0))) 
```



#### Share by Month Bar and Plot

```{r , echo=FALSE  ,  results = "asis", fig.align='center', warning=FALSE, message=FALSE, fig.width=9}
mnth <- prop.table(xtabs(
  cnt ~ month + brnd_nm,
  data = df[df$year == 2014 ,] ),1)

share <- as.data.frame(ftable(mnth))
share$Freq <- share$Freq*100

ggplot(share, aes(x = month, y = Freq, fill = brnd_nm))+ ylim(0,100) + 
  geom_bar(colour="black", stat="identity",   size=.3) +
 scale_fill_brewer(palette = 12) +
  #geom_text(aes(label = share), size = 6) +
  labs(title = "2014 Market Share by Brand inCounts", 
    y = "Share", x = "Month", fill = "Brand Name") +
    theme_bw() 
```


#### Trending the brands shows there is clearly inaccurate data in the last month.

```{r , echo=FALSE  ,  results = "asis", fig.align='center', warning=FALSE, message=FALSE, fig.width=9}
# trended
ggplot(share, aes(x = month, y = Freq, 
      group = brnd_nm,  colour = brnd_nm)) + geom_line(size = 2 ) +
   scale_fill_brewer(palette = 12) +
  labs(title = "2014 Market  Share by in Counts", 
    y = "Share", x = "Month", fill = "Brand Name") +
    theme_bw() 

# overlayed histograms
# quantile(df$cnt); quantile(df$cnt, .9)
```

#### Histogram of Counts in 2014 by Brand

```{r , echo=FALSE  ,  results = "asis", fig.align='center', warning=FALSE, fig.width=9, message=FALSE}
df %>% filter(year == 2014) %>%
ggplot(., aes(x = cnt, fill = brnd_nm)) +
  geom_histogram(bindwidth = 2, colour = "black") +
  facet_wrap(~ brnd_nm) +
  #geom_vline(aes(xintercept = mean(cnt, na.rm=T))
  #           , color = "red", linetype = "dashed")
  scale_fill_brewer(palette = 12) +
  xlim(0,40) +
   scale_fill_brewer(palette = 5) +
  labs(title = "2014 Market Trends by Brand in Counts", 
    y = "Share", x = "", fill = "Brand Name") +
    theme_bw() 
```

#### Counts in 2014 by Brand

```{r , echo=FALSE  ,  results = "asis", fig.align='center', warning=FALSE, fig.width=9, message=FALSE}
df %>% filter(year == 2014) %>%
  ggplot(., aes(x = brnd_nm , y = cnt, fill = brnd_nm)) + geom_boxplot() +
  guides(fill=FALSE) + ylim(0,40) +
  scale_fill_brewer(palette = 12) +
  coord_flip() +  xlab("Brand Name") + ylab("df Counts") + 
    ggtitle("Spread of Counts Below 95% Quantile") +  
    theme_bw()
```

#### Measuring outliers there are in counts
-- First is a general spread
-- Next is the 90th percentile 
-- Next is the 95th percentile
-- Finally the range of values is shown

```{r , echo=TRUE ,  results = "asis", fig.align='center', warning=FALSE,  message=FALSE, collapse=TRUE }
round(summary(df$cnt),2)
round(quantile(df$cnt, 0.95),2)
round(quantile(df$cnt, 0.90),2)
round(range(df$cnt),2)
```

#### Fitting a linear model to capture the the trend 

``` {r , echo=FALSE  , warning=FALSE,  results = "asis", fig.align='center', warning=FALSE, fig.width=9,cache.comments=FALSE, message=FALSE}
smth <- df %>% 
  filter( brnd_nm %in% c('A', 'T')) %>% 
    select(brnd_nm, date,  amt) %>% 
     arrange(date) %>%
      group_by (date, brnd_nm) %>%
   summarize(Amounts = sum(amt, na.rm=T)) 
  
ggplot(smth, aes(date, Amounts,  group = brnd_nm, colour = brnd_nm) ) + 
  geom_line( size = 1, aes(linetype = brnd_nm )) +
  geom_point(size = 2, fill ="white") +
   facet_grid(~ brnd_nm) +
  scale_fill_brewer(palette = 12) +
  geom_smooth() +
    xlab("Date") + ylab("Total Amount") + 
    ggtitle("Total Amount with 95% CI by Brand") +  
   theme(axis.text = element_text( size=.5)) +
    theme_bw()
````

####The Year Over Year Change Across Counts

``` {r  , echo=FALSE  ,  results = "asis", fig.align='center', warning=FALSE, fig.width=9}
load("yearly.RData")
load("brnd_lookup.RData")

yramounts <- left_join(yr_comp, brnd, by = c("BRNDID"))
yramounts <- tbl_df(yramounts)
names(yramounts) <- tolower(names(yramounts))
yramounts$brnd_nm <- toupper(yramounts$brnd_nm)

```

#### Examining the Y/Y change between Brands

``` {r  , echo=FALSE  ,  results = "asis", fig.align='center', warning=FALSE, message=FALSE, fig.width=9}
y_tbl <- yramounts %>% 
  select(year, month, cnt, brnd_nm) %>%
  group_by(year, month, brnd_nm) %>%
  summarize(ttl_cnt = sum(cnt, na.rm=T)) %>%
  dcast(., brnd_nm + month ~ year)

names(y_tbl) <- c("brnd_nm", "month", "year_2013", "year_2014")

y_tbl %>%
  mutate( diff = year_2014 / year_2013) %>%
  ggplot(., aes(x = factor(month), y = diff
                ,group = brnd_nm, colour = brnd_nm)) + 
  geom_line( size = 2) + ylim(0,140) +
   scale_fill_brewer(palette = 5) +
  labs(title = "Y/Y Index of Counts by Brands", 
    y = "Count Index", x = "", fill = "Brand Name") +
    theme_bw() 
rm(y_tbl)
````


####Trend of Counts Over Time, show steep increase for after launch.

``` {r , echo=FALSE  , warning=FALSE, results = "asis", fig.align='center', fig.width=9}
  df %>% 
  filter( brnd_nm == "T") %>%
  select(date, cnt) %>% 
  group_by (date) %>%
  summarize( Counts = sum(cnt, na.rm=T)) %>%
  ggvis( ~ factor(date) , ~Counts ) %>% layer_lines( stroke :="#0040FF" , strokeWidth := 4 ) %>% 
  add_axis("x", title="Period", title_offset = 50 ,  properties = axis_props(labels = list( fontSize = 8))) %>% 
  add_axis("y", title="") %>%
  scale_numeric("y", domain = c(0,  90000)) %>%
   add_axis("x", orient = "top", ticks = 0, title = "T Counts By Month",
           properties = axis_props(
             axis = list(stroke = "white"),
             labels = list(fontSize = 0))) 
````

####Trend of counts shows that stable market with one exception.

``` {r  , echo=FALSE  ,  results = "asis", fig.align='center', warning=FALSE, fig.width=10}
df %>% 
  select(brnd_nm, date,  cnt) %>% 
  group_by (date, brnd_nm) %>%
  summarize( Counts = sum(cnt, na.rm=T)) %>%
ggplot(. , aes(x = factor(date), y = Counts,
               group = brnd_nm,  colour = brnd_nm)) + geom_line(size = 2 ) +
   scale_fill_brewer(palette = 12) +
  labs(title = "Brands : Counts By Month" ,
    y = "Counts", x = "Date") +
  theme(axis.text=element_text(size=6), axis.title=element_text(size=6)) +
    theme_bw() 

````


####Overall Counts and Amounts by the Brand for 2014

``` {r  , echo=FALSE  ,  results = "asis", fig.align='center', warning=FALSE, fig.width=9, comment=F, message=F}
 df %>% 
  filter( year == 2014 ) %>%
  ggplot(., aes(x = amt , y = cnt, group = brnd_nm, colour = brnd_nm)) + geom_point(shape=1) +
  xlim(0,6000) + ylim(0, 20) + 
  facet_wrap(~ brnd_nm) +
  scale_fill_brewer(palette = 12) +
  geom_smooth()+
  labs(title = "Count and Amounts in 2014 by Brand" ,  y = "Count", x = "Amount") +
  theme_bw() 
````

